---
title: "Exercise 4"
author: '[Your first name] [Your last name]'
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: true
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

**For this week's assignment, you'll construct a period multiple decrement table and interpret the results. For next week's assignment, using the same dataset, you'll apply your skills at building associated decrement life tables (including cause-deleted life tables), and conducting cause-specific decomposition of life expectancy differences.**

## Setup {.unnumbered}

```{r}
if (!require("pacman")) install.packages("pacman")
# p_load function loads packages if installed, or install then loads otherwise
pacman::p_load(dplyr, ggplot2, knitr, kableExtra, readr)
```

## Load data {.unnumbered}

```{r}
rus_mxi <- readRDS(url("https://github.com/hanowell/uwsoc533a-exercise-05/blob/main/data/rus_mxi.rds?raw=true"))
```

## About the data {.unnumbered}

**You've just loaded age- and cause-specific mortality rates for **